ESI for ``Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers''

### 2018-08-15 reports/paper.tex
third draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This is the LaTeX ARTICLE template for RSC journals
%Copyright The Royal Society of Chemistry 2016
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\includegraphics{head_foot/dates} & \noindent\normalsize{Deep eutectic solvents present a novel class of non-aqueous room temperature solvent with tunable properties, that are capable of promoting the self-assembly of surfactant molecules. However, the solvation model in these systems still challenges the classic understanding of amphiphilicity. In this work, we present the first example of the self-assembly of phospholipid monolayers at the interface between air and a non-aqueous solvent. Furthermore, we use novel, chemically-consistant Bayesian modelling of X-ray and neutron reflectometry measurements to show the ability of the deep eutectic solvent to interact with the phosphatidylglycerol lipid head group, leading to an apparent increase in the component volume compared to that observed in water. No such change was observed for the phosphocholine head group, indicating that the interaction is head group specific.} \\

@@

%Please use \dag to cite the ESI in the main text of the article.
%If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
\footnotetext{\dag~Electronic Supplementary Information (ESI) available: A series Jupyter notebooks and Python plotting scripts, allowing for a fully reproducible analysis of all data presented herein as well as figures showing the probability distribution functions for each of the lipid models. See DOI: 10.1039/b000000x/}
%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required

@@

The formation of phospholipid monolayers at the air/liquid interface also plays a key role in many biological and technological processes. The solvent-specific solubility of different components of the phospholipid results in the formation of a stable monolayer of phospholipid at the interface.\cite{Mohwald1990} Phospholipids contain a charged headgroup, either anionic or zwitterionic, and investigations at the air-salt water interface have revealed the importance of the phospholipid-ion interactions on structure, monomer packing, and stability of the monolayer.\cite{Mohwald1990,Kewalramani2010} Despite the broad interest in these systems, the presence of stable phospholipid monolayers at the interface between air and a non-aqueous media has not been previously reported, to the best of the authors' knowledge.

@@

In this work, we present the first investigation of the structure of phospholipid monolayers at the air-DES interface, as determined by chemically-consistant modelling of X-ray reflectometry (XRR) measurements. Four different phospholipids; 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (DMPG), were studied at the interface between a 1:2 mixture of choline chloride:glycerol and air. This allowed the nature of two, chemically distinct, phospholipid headgroups to be understood in this non-aqueous solvent, in addition to the effect of the tail chain length. The analysis was then extended to model complementary neutron reflectometry measurements for two contrasts of DMPC and DPPC at a single surface pressure.

@@

DPPC (> 99 \%, C$_{16}$ tails), DMPC (> 99 \%, C$_{14}$ tails), and DMPG (> 99 \%, C$_{14}$ tails) were supplied by Avanti Polar Lipids and DLPC (> 99 \%, C$_{12}$ tails) was supplied by Sigma-Aldrich and all were used as received. Deuterated versions of DPPC (d$_{62}$-DPPC, > 99 \%, deuterated tails-only) and DMPC (d$_{54}$-DPPC, > 99 \%, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification. These phospholipids were dissolved in chloroform (0.5 mg/mL) at room temperature. PC indicates the molcule contains a phosphocholine head group, where PG contains a phosphoditylglycerol head group, these are shown in Figure \ref{fig:heads}.

\begin{figure}
	\centering
	\includegraphics[width=0.30\textwidth]{figures/head_groups}
	\caption{The two lipid classes with different head groups compared in this study, where R indicates the hydrocarbon tail; (a) phosphatidylglycerol, (b) phosphocholine. Figure files are available under MIT License.\cite{mccluskey_2018}}
	\label{fig:heads}
\end{figure}

\begin{table*}
	\small
	\caption{\ Lipid component volumes extracted from different literature sources. $V_l$ corresponds to the total lipid volume, MD to molecular dynamics simulation, WAXS to wide-angle X-ray scattering, NB to neutral buoyancy and DVTD to differential vibrating tube densimetry}
	\label{tab:water}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllllll}
		\hline 
		Lipid & DPPC & & & & DMPC & & & DMPG \\
		\hline
		Reference & [\citen{Armen1998}] & [\citen{Petrache1997}] & [\citen{Sun1994}] & [\citen{Tardieu1973}] & [\citen{Kucerka2004}] & [\citen{Nagle1978}] & [\citen{Schmidt1985}] & [\citen{Pan2012}] \\
		$V_l$/\AA$^3$ &1216.96 & 1219 & 1148 & 1224 & 1101 & 1061 & 1094 & 1058 \\
		$V_h$/\AA$^3$ & 326.00 & 324 & 319 & 360 & 319 & 344 & & 291 \\
		$V_t$/\AA$^3$ & 890.96 & 895 & 829 & 864 & 782 & 717 & & 767 \\
		Method & MD & MD & WAXS & NB & NB & NB & DVTD & DVTD \\
		T/$^\circ$C & 25 & 50 & 24 & 25 & 30 & 30 & & 30 \\
		\hline
	\end{tabular*}
\end{table*}
In the XRR experiment, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration. After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface. Surface concentration was modified by closing and opening the PTFE barriers of a Langmuir trough. In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a managageable level) it was not possible to use a Langmuir trough. Instead, small Delrin adsorption troughs were used that did not have controlable barriers. This resulted in no control over the surface pressure of the measurement and therefore it was not possible to co-refine NR contrasts.

\subsection{Methods}
XRR measurements were taken on I07 at Diamond Light Source, at 12.5 keV photon energy using the double-crystal-deflector.\cite{Arnold2012} The reflected intensity was measured in a momentum transfer range from 0.018 to 0.7 \AA$^{-1}$. The data were normalised with respect to the incident beam and the background was measured from off-specular reflection and subsequently subtracted. Samples were equilibrated for at least one hour and preserved under an argon atmosphere to minimise the adsorption of water by the subphase. XRR data were collected for each of the lipids, DMPC, DPPC, DLPC and DMPG at two surface pressures, 20 mNm$^{-1}$ and 30 mNm$^{-1}$, as measured with an aluminium Wilhelmy plate; all measurements were made at 22 $^\circ$C. The aluminium Wilhelmy plate was used over a traditional paper plate due to the low wetability of paper by the DES.

The NR experiments were performed on FIGARO at the Institut Laue-Langevin using the time-of-flight method.\cite{Campbell2011} Data at two incident angles of 0.62$^\circ$ and 3.8$^\circ$ were measured to provide a momentum transfer range from 0.005 to 0.18 \AA$^{-1}$. A single surface concentration for each system and contrast was measured. Similar to the X-ray procedure, samples were given enough time to equilibrate (at least two hours) and kept under an inert atmosphere.

\subsection{Data analysis}
The use of XRR and NR to analyse the structure of phospholipids on the surface of water has a history extending over many years.\cite{Mohwald1990,Kewalramani2010,Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990} The models used in the rationalisation of XRR and NR data have varied significantly in number of layers present, use of interfacial roughness, and parameterisation of physical constraints. Frequently these physical constraints include the component volumes of the phospholipid head and tail groups, with values taken from other techniques, such as those shown in Table \ref{tab:water}. Additionally, a recent evaluation of the applicability of different models to surfactant and phospholipid monolayers from the NR perspective has been published,\cite{Campbell2018} suggests possible oversights in the modelling of NR data.

In Table \ref{tab:water}, there appears to be a general consensus that the component volume of the phosphocholine (PC) headgroup is in the range from 320 \AA$^3$ to 360 \AA$^3$ while data from a single source give the phosphatidylglycerol (PG) headgroup as 291 \AA$^3$. However, we do not know whether the head group component volumes used in the literature, that are derived from water-based measurements, will be appropriate for this work, which involves a non-aqueous solvent. The charged nature of the zwitterionic and anionic lipid heads means that they are likely to have different interactions with polar, but neutral, water as compared to the charged DES.\cite{Sanchez-Fernandez2018} Further it has been shown that in the liquid condensed phase, which is often present at high surface pressure, the lipid tails will become compressed resulting in a component volume less than measured in the liquid expanded phase,\cite{Marsh2010,Small1984} and the need to take into account the compaction of the hydrocarbon chains of phospholipid monolayers according to their phase in the modelling of NR data has recently been demonstrated.\cite{Campbell2018}

To allow for the use of a chemically-consistant model, where the lipid component volumes were allowed to vary, the Python library refnx\cite{Nelson2018} as used. This software allows the inclusion of a custom model from which the parameters fed into the Abel\`{e}s model for the reflection of light at a given number of stratified interfaces,\cite{Abeles1950,Parratt1954} that is typical for reflectometry fitting, are obtained. This custom model, along with a series of Jupyter notebooks showing, in full, the analysis performed, can be found in the ESI and is available under an MIT license.\cite{mccluskey_2018}

This chemically-consistent model involves two layers consisting of head groups at the interface with the solvent and tail groups at the interface with the air, and is based on the assumption that at the surface pressures measured the monolayer tail will be in the condensed phase. The head groups have a calculated scattering length, $b_h$, (found as a summation of the number of electrons in the head group multiplied by the classical radius of the electron for XRR or the summation of the neutron scattering length for NR), and a component volume, $V_h$. These head groups make up a layer with a given thickness, $d_h$, and roughness, $\sigma_h$, within which some volume fraction of solvent can intercalate, $\phi_h$. The tail groups also have a calculated scattering length, $b_t$, and a component volume, $V_t$, however the thickness of the tail group layer, $d_t$, is found from the length of the carbon tail, $t_t$, and angle that the chain is tilted by with respect to the interface normal, $\theta_t$, 
\begin{equation}
\label{equ:tl}
d_t = t_t \cos{\theta_t}.
\end{equation}
The scattering length density (SLD) of the tail and head layers used in the Abel\`{e}s model can therefore be found as follows, 
\begin{equation}
\text{SLD}_i = \frac{b_i}{V_i}(1 - \phi_i) + \text{SLD}_{s}\phi_i,
\end{equation}
where $\text{SLD}_{s}$ is the scattering length density of the subphase (DES), and $i$ indicates either the tail or head layer, and it is assumed that the tail layer contains no solvent or air, e.g. $\phi_t = 0$. To ensure that the number density of head groups and pairs of tail groups is the same, the following constraint was included in the model,\cite{Braun2017}
\begin{equation}
\label{equ:ht}
d_h = \frac{V_hd_t}{V_t(1-\phi_h)}. 
\end{equation}
A single value for the interfacial roughness was fitted for all interfaces, as there is only a single lipid molecule type in each monolayer. Therefore, any capillary wave roughness at the air-DES interface is carried equally through the layers.

In the first of two steps, this custom model was used to co-refine the component volume of the lipid head group, $V_h$, and the volume of the tail group, $V_t$, across the XRR measurments at two surface concentrations. The following parameters were allowed to vary; $\theta_t$, $d_h$, $\phi_h$, and $\sigma_{t,h,s}$, independently across the two surface pressures, while others, shown in Table \ref{tab:invariant}, were held constant at the values given. The length of the carbon chain was kept constant to the value determined by the Tanford equation,\cite{Tanford1980} this is valid due to the condensed nature of the monolayer at this surface concentration resulting in the extended, staggered conformation of the chain being likely. For each co-refinement of two XRR measurements, there were, in total, eight degrees of freedom in the fitting process.

\begin{table}[h]
	\small
	\caption{\ The invariant parameters within the chemically-sensible model. 
	$^a$Values obtained from the Tanford formula.\cite{Tanford1980} $^b$Values extracted from Sanchez-Fernandez \emph{et al.}\cite{Sanchez-Fernandez2016}}
	\label{tab:invariant}
	\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}lllll}
		\hline
		Component & $b_t$/fm & $b_h$/fm & $t_t$/\AA & $\text{SLD}$/$\times10^{-6}$\AA$^{-2}$ \\
		\hline
		X-ray & & & & \\
		DLPC & 5073 & 4674 & 15.5$^a$ & -- \\
		DMPC & 5985 & 4674 & 18.0$^a$ & -- \\
		DPPC & 6897 & 4674 & 20.5$^a$ & -- \\
		DMPG & 5985 & 4731 & 18.0$^a$ & --\\
		Air & -- & -- & -- & 0\\
		DES & -- & -- & -- & 10.8$^b$ \\
		\hline
		Neutron & & & & \\
		d$_{54}$-DMPC & 5329.8 & 602.7 & 18.0$^a$ & -- \\
		d$_{62}$-DPPC & 6129.2 & 602.7 & 20.5$^a$ & -- \\
		h-DES & -- & -- & -- & 0.43$^b$  \\
		hd-DES & -- & -- & -- & 3.15$^b$ \\
		\hline
	\end{tabular*}
\end{table}

In the second step, the head group and tail group component volumes determined from XRR were used in the refinement of the custom model against the NR measurements. Due to the lack of contrast present between the hydrogenous PC head group and the solvent, it was necessary to also constrain the thickness of the head layer based on that determined from the lowest surface concentration XRR measurement. Therefore, a different constraint was required to ensure parity between the numbers of head group and pairs of tail groups,
\begin{equation}
\label{equ:phih}
\phi_h =  1 - \bigg(\frac{d_tV_h}{V_td_h}\bigg).
\end{equation}
Table \ref{tab:invariant} gives the details of the scattering length and SLDs used as invariant parameters in the custom model, while the tail chain lengths were the same as for the XRR. The definition of the head and tail component volumes and the head thickness allowed the number of variable parameters to be reduced to two, namely the chain tilt angle, $\theta_t$, and the interfacial roughness, $\sigma_{t,h,s}$.

In both cases, the refinement of the custom model to the experimental data involved the transformation of the reflectometry calculated from the model and the data into $Rq^4$ such that the contribution of the Fresnel decay was removed, before using the differential evoluation method available to refnx from the scipy library,\cite{Jones2001} to find the parameters that gave the best fit to the data. The parameter space was then probed using the MCMC method available through emcee, which allowed for an estimate of the PDF associated with each parameter. In the MCMC sampling, 200 walkers were used over 1000 iterations, following an equilibration of 200 iterations.

\section{Results \& Discussion}
\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{figures/DLPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DPPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPG_all_data}
	\caption{The XRR profiles, SLD profiles and PDFs of the head component volume for each of the four lipids, the lower surface concentration is shown in blue while the higher in green; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. The different surface pressure XRR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by $5\times10^{-6}$ \AA$^{-2}$, for clarity. Figure files are available under MIT License.\cite{mccluskey_2018}}
	\label{fig:lipids}
\end{figure}
The chemically-consistent model was co-refined across the two surface pressure XRR measurements for each lipid. The resulting XRR profiles, associated SLD profiles and the PDF for the head group component volumes are shown in Figure \ref{fig:lipids}. Table \ref{tab:liptab} gives details of all varying parameters for each system as well as the details of the $d_t$ and $d_h$ which are determined from Eqns. \ref{equ:tl} and \ref{equ:ht} respectively, each value is given with asymmetric uncertainies that corresponds to a 95 \% confidence interval of the PDF; the full PDF plots can be found in the ESI. 
\begin{table*}
	\small
	\caption{\ The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the XRR models. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}. The values of $d_h$ were obtained from the appropriate use of Eqn. \ref{equ:ht}}
	\label{tab:liptab}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllllll}
		\hline
		Lipid & DLPC & & DMPC & & DPPC & & DMPG & \\
		Surface Pressure/mNm${-1}$ & 20 & 30 & 20 & 30 & 20 & 30 & 20 & 30 \\
		\hline
		$\theta_t$/$^\circ$ & \input{../output/dlpc/angle4.txt} & \input{../output/dlpc/angle5.txt} & \input{../output/dmpc/angle4.txt} & \input{../output/dmpc/angle5.txt} & \input{../output/dppc/angle4.txt} & \input{../output/dppc/angle5.txt} & \input{../output/dmpg/angle4.txt} & \input{../output/dmpg/angle5.txt} \\
		$V_t$/\AA$^3$ & \input{../output/dlpc/vt.txt} & \input{../output/dlpc/vt.txt} & \input{../output/dmpc/vt.txt} & \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} & \input{../output/dppc/vt.txt} & \input{../output/dmpg/vt.txt} & \input{../output/dmpg/vt.txt} \\

>>>>

%%\%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This is the LaTeX ARTICLE template for RSC journals
%Copyright The Royal Society of Chemistry 2016
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\includegraphics{head_foot/dates} & \noindent\normalsize{Deep eutectic solvents present a novel class of non-aqueous room temperature solvent with tunable properties, that are capable of promoting the self-assembly of surfactant molecules. However, the solvation model in these systems still challenges the classic understanding of amphiphilicity. In this work, we present the first example of the self-assembly of phospholipid monolayers at the interface between air and a non-aqueous solvent. Furthermore, we use novel, chemically-consistant Bayesian modelling of X-ray and neutron reflectometry measurements to show the ability of the deep eutectic solvent to interact with the phosphatidylglycerol lipid head component, leading to an apparent increase in the component volume compared to that observed in water. No such change was observed for the phosphocholine head component, indicating that the interaction is head component specific.} \\

@@

%Please use \dag to cite the ESI in the main text of the article.
%If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
\footnotetext{\dag~Electronic Supplementary Information (ESI) available: All datasets, figure file and analysis/plotting scripts, allowing for a fully reproducible analysis of all work presented herein. See DOI: 10.6084/m9.figshare.6661784}
%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required

@@

The formation of phospholipid monolayers at the air/liquid interface also plays a key role in many biological and technological processes. The solvent-specific solubility of different components of the phospholipid results in the formation of a stable monolayer of phospholipid at the interface.\cite{Mohwald1990} Phospholipids contain a charged head, either anionic or zwitterionic, and investigations at the air-salt water interface have revealed the importance of the phospholipid-ion interactions on structure, monomer packing, and stability of the monolayer.\cite{Mohwald1990,Kewalramani2010} Despite the broad interest in these systems, the presence of stable phospholipid monolayers at the interface between air and a non-aqueous media has not been previously reported, to the best of the authors' knowledge.

@@

In this work, we present the first investigation of the structure of phospholipid monolayers at the air-DES interface, as determined by chemically-consistant modelling of X-ray reflectometry (XRR) measurements. Four different phospholipids; 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (DMPG), were studied at the interface between a 1:2 mixture of choline chloride:glycerol and air. This allowed the nature of two, chemically distinct, phospholipid head components to be understood in this non-aqueous solvent, in addition to the effect of the tail chain length. The analysis was then extended to model complementary neutron reflectometry measurements for two contrasts of DMPC and DPPC at a single surface pressure.

@@

DPPC (> 99 \%, C$_{16}$ tails), DMPC (> 99 \%, C$_{14}$ tails), and DMPG (> 99 \%, C$_{14}$ tails) were supplied by Avanti Polar Lipids and DLPC (> 99 \%, C$_{12}$ tails) was supplied by Sigma-Aldrich and all were used as received. Deuterated versions of DPPC (d$_{62}$-DPPC, > 99 \%, deuterated tails-only) and DMPC (d$_{54}$-DPPC, > 99 \%, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification. These phospholipids were dissolved in chloroform (0.5 mg/mL) at room temperature. PC indicates the molcule contains a phosphocholine head component, where PG contains a phosphoditylglycerol head component, these are shown in Figure \ref{fig:heads}.

\begin{figure}
	\centering
	\includegraphics[width=0.30\textwidth]{figures/head_groups}
	\caption{The two lipid classes with different head components compared in this study, where R indicates the hydrocarbon tail; (a) phosphatidylglycerol, (b) phosphocholine. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:heads}
\end{figure}

\begin{table*}
	\small
	\caption{\ Lipid component volumes extracted from different literature sources. $V_l$ corresponds to the total lipid volume, MD to molecular dynamics simulation, WAXS to wide-angle X-ray scattering, NB to neutral buoyancy and DVTD to differential vibrating tube densimetry. $^a$ The values for the head component in Kucerka \emph{et al.},\cite{Kucerka2004} were taken from Balgav\'{y} \emph{et al}.\cite{Balgavy2001}}
	\label{tab:water}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}llllllllll}
		\hline
    Lipid & DPPC & & & DMPC & & DLPC & & DMPG & POPG \\
    \hline
    Reference & [\citen{Armen1998}] & [\citen{Sun1994}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Armen1998}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Armen1998}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Pan2012}] & [\citen{Kucerka2012}] \\
    \hline
    $V_l$/\AA$^3$ & $1287.3\pm25.5$ & $1148\pm2$ & $1264.2\pm32.1$ & $1172.5\pm25.1£$ & $1155.4\pm30.0£$ & $1057.7\pm24.7$ & $1046.6\pm28.0$ & $1011.4$ & $1203$ \\
    $V_t$/\AA$^3$ & $966.4\pm5.4$ & $829\pm4$ & $924.7\pm17.6$ & $851.5\pm5.0£$ & $815.9\pm15.5£$ & $736.8\pm4.6$ & $707.1\pm13.5$ & $720.4$ & $914$ \\
    $V_h$/\AA$^3$ & $320.9\pm20.1$ & $319\pm6$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $291.0$ & $289$ \\
    Method & MD & WAXS & NB & MD & NB & MD & NB & DVTD & MD \\
    T/$^\circ$C & 50 & 24 & 30 & 50 & 30 & 50 & 30 & 20 & 25 \\
	\end{tabular*}
\end{table*}
In the XRR experiment, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration. After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface. Surface concentration was modified by closing and opening the PTFE barriers of a Langmuir trough. In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a managageable level) it was not possible to use a Langmuir trough. Instead, small Delrin adsorption troughs were used that did not have controlable barriers. This resulted in no control over the surface pressure of the measurement and therefore it was not possible to co-refine NR contrasts.

\subsection{Methods}
XRR measurements were taken on I07 at Diamond Light Source, at 12.5 keV photon energy using the double-crystal-deflector.\cite{Arnold2012} The reflected intensity was measured in a momentum transfer range from 0.018 to 0.7 \AA$^{-1}$. The data were normalised with respect to the incident beam and the background was measured from off-specular reflection and subsequently subtracted. Samples were equilibrated for at least one hour and preserved under an argon atmosphere to minimise the adsorption of water by the subphase. XRR data were collected for each of the lipids, DLPC, DMPC, DPPC and DMPG at four surface pressures (DLPC: 20, 25, 30, and 35 mNm$^{-1}$, DMPC: 20, 25, 30, and 40 mNm$^{-1}$, DPPC: 15, 20, 25, and 30 mNm$^{-1}$, DMPG: 15, 20, 25, and 30 mNm$^{-1}$), as measured with an aluminium Wilhelmy plate; all measurements were conducted at 22 $^\circ$C. The aluminium Wilhelmy plate was used over a traditional paper plate due to the low wetability of paper by the DES.

The NR experiments were performed on FIGARO at the Institut Laue-Langevin using the time-of-flight method.\cite{Campbell2011} Data at two incident angles of 0.62$^\circ$ and 3.8$^\circ$ were measured to provide a momentum transfer range from 0.005 to 0.18 \AA$^{-1}$. Two surface pressures for each system and contrast was measured (DMPC: 20 and 25 mNm$^{-1}$, DPPC: 15 and 20 mNm$^{-1}$). Similar to the X-ray procedure, samples were given enough time to equilibrate (at least two hours), kept under an inert atmosphere, and all measurements were conducted at 22 $^\circ$C.

\subsection{Data analysis}
The use of XRR and NR to analyse the structure of phospholipids on the surface of water has a history extending over many years.\cite{Mohwald1990,Kewalramani2010,Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990} The models used in the rationalisation of XRR and NR data have varied significantly in number of layers present, use of interfacial roughness, and parameterisation of physical constraints. Frequently these physical constraints include the component volumes of the phospholipid head and tail components, using values taken from other techniques, such as those shown in Table \ref{tab:water}. Additionally, a recent evaluation of the applicability of different models to surfactant and phospholipid monolayers from the NR perspective has been published,\cite{Campbell2018} suggests possible oversights in the modelling of NR data.

In Table \ref{tab:water}, there appears to be a general consensus that the component volume of the phosphocholine (PC) head is in the range from 320-360 \AA$^3$ while the phosphatidylglycerol (PG) head is in the range 289-291 \AA$^3$. However, we do not know whether the head component volumes used in the literature, that are derived from water-based measurements, will be appropriate for this work, which involves a non-aqueous solvent. The charged nature of the zwitterionic and anionic lipid heads means that they are likely to have different interactions with polar, but neutral, water as compared to the charged DES.\cite{Sanchez-Fernandez2018} Furthermore, it has been shown that in the liquid expanded and condensed phases, which are often present at high surface pressure, the lipid tails will become compressed resulting in a component volume less than measured from other techniques,\cite{Marsh2010,Small1984} and the need to take into account the compaction of the hydrocarbon chains of phospholipid monolayers according to their phase in the modelling of NR data has recently been demonstrated.\cite{Campbell2018}

To allow for the use of a chemically-consistant model, where the lipid component volumes were allowed to vary, the Python library refnx\cite{Nelson2018} as used. This software allows the inclusion of a custom model from which the parameters fed into the Abel\`{e}s model for the reflection of light at a given number of stratified interfaces,\cite{Abeles1950,Parratt1954} that is typical for reflectometry fitting, are obtained. This custom model, along with a series of Jupyter notebooks showing, in full, the analysis performed, can be found in the ESI and is available under CC-BY.\cite{mccluskey_2018}

This chemically-consistent model involves two layers; the head layer at the interface with the solvent and tail layer at the interface with the air. The head components have a calculated scattering length, $b_h$, (found as a summation of the number of electrons in the head component multiplied by the classical radius of the electron for XRR or the summation of the neutron scattering lengths for NR), and a component volume, $V_h$. These head components make up a layer with a given thickness, $d_h$, and roughness, $\sigma_h$, within which some volume fraction of solvent can intercalate, $\phi_h$. The tail components also have a calculated scattering length, $b_t$, and a component volume, $V_t$. In this work, the thickness of the tail layer, $d_t$, was found from the length of the carbon tail, $t_t$, and angle that the chain is tilted by with respect to the interface normal, $\theta_t$,
%
\begin{equation}
\label{equ:tl}
d_t = t_t \cos{\theta_t}.
\end{equation}
%
The was used to impose a chemically-sensible maximum on the thickness of the lipid tail layer, e.g. it cannot be thicker than the maximum extended length of the lipid tail. The scattering length density (SLD) of the tail and head layers used in the Abel\`{e}s model can therefore be found as follows,
%
\begin{equation}
\text{SLD}_i = \frac{b_i}{V_i}(\phi_i) + \text{SLD}_{s}(1-\phi_i),
\end{equation}
%
where, $\text{SLD}_{s}$ is the scattering length density of the subphase (DES), and $i$ indicates either the tail or head layer; it is assumed that the tail layer contains either solvent nor air, e.g. $\phi_t = 0$. To ensure that the number density of head components and pairs of tail components is the same, the following constraint was included in the model,\cite{Braun2017}
%
\begin{equation}
\label{equ:phih}
\phi_h =  1 - \bigg(\frac{d_tV_h}{V_td_h}\bigg).
\end{equation}
%
Based on the work of Campbell \emph{et al.},\cite{Campbell2018} a single value for the interfacial roughness was fitted for all interfaces, as there is only a single lipid molecule type in each monolayer. Therefore, any capillary wave roughness at the air-DES interface is carried equally through the layers.

In the first of two steps, this custom model was used to co-refine the component volume of the lipid head component, $V_h$, the volume of the tail component, $V_t$, and the head thickness, $d_h$ across the XRR measurments at four surface concentrations. The head thickness was not permitted to vary with surface pressure to reduce the dimensionality of the parameter space, this based on the understanding that the increasing surface pressure should not significantly influence the head layer thickness. The following parameters were allowed to vary; $\theta_t$, and $\sigma_{t,h,s}$, independently across the surface pressures, while others, shown in Table \ref{tab:invariant}, were held constant at the values given. The length of the carbon chain was kept constant at the value determined by the Tanford equation,\cite{Tanford1980} due to the nature of the monolayer at the surface pressures resulting in the extended, staggered conformation of the chain being likely. For each co-refinement of four XRR measurements, there were, in total, eleven degrees of freedom in the fitting process.

%
\begin{table}[h]
	\small
	\caption{\ The invariant parameters within the chemically-consistent model.
	$^a$Values obtained from the Tanford formula.\cite{Tanford1980} $^b$Values obtained from Sanchez-Fernandez \emph{et al.}\cite{Sanchez-Fernandez2016}}
	\label{tab:invariant}
	\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}lllll}
		\hline
		Component & $b_t$/fm & $b_h$/fm & $t_t$/\AA & $\text{SLD}$/$\times10^{-6}$\AA$^{-2}$ \\
		\hline
		X-ray & & & & \\
		DLPC & 5073 & 4674 & 15.5$^a$ & -- \\
		DMPC & 5985 & 4674 & 18.0$^a$ & -- \\
		DPPC & 6897 & 4674 & 20.5$^a$ & -- \\
		DMPG & 5985 & 4731 & 18.0$^a$ & --\\
		Air & -- & -- & -- & 0\\
		DES & -- & -- & -- & 10.8$^b$ \\
		\hline
		Neutron & & & & \\
		d$_{54}$-DMPC & 5329.8 & 602.7 & 18.0$^a$ & -- \\
		d$_{62}$-DPPC & 6129.2 & 602.7 & 20.5$^a$ & -- \\
		h-DES & -- & -- & -- & 0.43$^b$  \\
		hd-DES & -- & -- & -- & 3.15$^b$ \\
		\hline
	\end{tabular*}
\end{table}
%
In the second step, the head and tail component volumes, and head layer thickness determined from XRR were used in the refinement of the custom model against the NR measurements. Table \ref{tab:invariant} gives the details of the scattering lengths and SLDs used as invariant parameters in the custom model, while the tail chain lengths were the same as for the XRR. The definition of the head and tail component volumes, and the head thickness allowed the number of variable parameters to be reduced to two, namely the chain tilt angle, $\theta_t$, and the interfacial roughness, $\sigma_{t,h,s}$, for the co-refinement of two datasets.

In both cases, the refinement of the custom model to the experimental data involved the transformation of the reflectometry calculated from the model and the data into $Rq^4$ such that the contribution of the Fresnel decay was removed, before using the differential evoluation method available to refnx from the scipy library,\cite{Jones2001} to find the parameters that gave the best fit to the data. The parameter space was then probed using the MCMC method available through emcee,\cite{Foreman-Mackey2013} which allowed for an estimate of the PDF associated with each parameter. In the MCMC sampling, 200 walkers were used over 1000 iterations, following an equilibration of 200 iterations. The use of MCMC sampling allowed for the Bayesian inference of the probability density function for each of the variables and their respective interactions (shown in detail in the ESI). This allowed asymmetric confidence intervals to be found for each of the variable parameters. However, it is important to note that these are not \emph{true} confidence intervals, and account only for the uncertainty present in the data, e.g. they do no account of systematic uncertainty in the measurement that are underrepresented, or unrepresented, in the experimental dataset.
